Imputation of Incomplete Multilevel Data with \proglang{R}

### 2023-11-14 Manuscript/Imputation_of_Incomplete_Multilevel_Data.Rmd
update case III

```{r, include=FALSE, eval=FALSE}
library(here)
source(here("heckman","ipdfunction.R"),local =knitr::knit_global())
load(here("heckman","Obesity.rda"))
```

@@

```{r, dpi = 300,out.width="70%", fig.cap="Missing pattern", eval=FALSE}
library(ggpubr)
myplots <- lapply(1:5, function(i) {
  ggmice::plot_pattern(setDT(Obesity)[Cluster==i])+
  ggplot2::ggtitle(paste0("Cluster", i))
 })
ggarrange(myplots[[1]], myplots[[3]], nrow=1,common.legend = TRUE, legend="bottom")
```

@@

```{r obesity-icc, eval=FALSE}
Nulmodel <- lme4::lmer(BMI ~ 1 + (1|Cluster), data = Obesity)
#Performance::ICC(Nulmodel)
```

@@

```{r obesitymeth, eval=FALSE}
meth_mar <- find.defaultMethod(Obesity, ind.clust=1, I.small = 7,
                               ni.small = 100, prop.small = 0.4)
meth_mar["BMI"]<- "~ I(Weight / (Height)^2)"
meth_mar["Age"]<-"2l.2stage.pmm" 
```

For these imputation models, it is necessary to specify the prediction matrix, with the cluster
variable labelled as -2 and the predictor variable measured within clusters labelled as 2, encompassing all variables. We need to suprime the variable Time as this variable is not specified in the main model. We also  modify the relationship between BMI, weight and height in the prediction matrix to avoid circular predictions. Then we proceed to run the imputation model.
```{r obesity-predmar,out.width="70%", eval=FALSE}
pred_mar <- quickpred(Obesity)   # predictor matrix
pred_mar[,"Cluster"] <- -2 # clustering variable
pred_mar[,"Time"] <- 0
pred_mar[pred_mar==1] <- 2
pred_mar[c("Height", "Weight"), "BMI"] <- 0
ggmice::plot_pred(pred_mar)
imp_mar <- mice(data = Obesity, meth = meth_mar, pred = pred_mar,
                m=10, seed = 123, printFlag = FALSE)
```

```{r obesity-predmar1,eval=FALSE}
summary(complete(imp_mar,"long")$Weight)
```
We are also contemplating the utilisation of the ???pmm??? option, as the values imputed using a fully parametric method may be implausibly low for some patients.

```{r obesity-predmar_pmm, eval=FALSE}
meth_mar["Weight"]<-"2l.2stage.pmm" 
imp_mar_pmm <- mice(data = Obesity, meth = meth_mar, pred = pred_mar,
                    m=10, seed = 123, printFlag = FALSE)
```

```{r obesity-predmar_pmm1,eval=FALSE}
summary(complete(imp_mar_pmm,"long")$Weight)
#ggmice::ggplot_trace(imp_mar_pmm, "Weight")
```

@@

```{r obesity-predmnar, eval=FALSE}
pred_mnar <- pred_mar
pred_mnar["Weight","Time"]<- -3
#ggmice::plot_pred(pred_mnar)
```

@@

```{r obesity-predmnar1, eval=FALSE}
meth_mnar <- meth_mar
meth_mnar["Weight"]<- "3l.2stage.heckman"
```
Then we proceed to run the imputation model as before, after executing these imputation procedures, it is essential to assess convergence and the coherence of the imputed values.
```{r obesity-predmnar2,eval=FALSE}
imp_mnar<- mice(data = Obesity, meth = meth_mnar, pred = pred_mnar,
                m=10, seed = 123, printFlag = FALSE)
summary(complete(imp_mnar,"long")$Weight)
```

@@

```{r obesity-predmnar3,eval=FALSE}
summary(complete(imp_mnar,"long")$Weight)
```
Consequently, as before we  use the 'pmm', option but this time for the Heckman imputation, this approach ensures that the imputed values remain within the range of observable values. We then run the imputation model but this time using the option of pmm, to assure that weight values are in the range of the observable data, this can be implemented by setting the pmm parameter to true.

```{r obesity-predmnarp, eval=FALSE}
imp_mnar_pmm <- mice(data = Obesity, meth = meth_mnar, pred = pred_mnar,
                     m=10, seed = 123, pmm = T,  printFlag = FALSE)
```

We check the convergency of the results
```{r obesity-predmnarp1, eval=FALSE}
summary(complete(imp_mnar_pmm,"long")$Weight)
#ggmice::ggplot_trace(imp_mar_pmm, "Weight")
```

@@

```{r models, eval=FALSE}
library(ggplot2)
cc_rs<- with(setDT(Obesity)[complete.cases(Obesity),],
             lme( BMI ~ Age + FamOb + Gender, random=~1+Age|Cluster))
mar_rs <- with(imp_mar,lme( BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
mar_pmm_rs <- with(imp_mar_pmm,lme( BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
mnar_rs<- with(imp_mnar,lme(BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
mnar_pmm_rs<- with(imp_mnar_pmm, lme(BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
list_models<-list(cc_rs,mar_rs,mar_pmm_rs,mnar_rs,mnar_pmm_rs)
plot_models(list_models,
            mod_name = c("Complete case", "MAR","MAR_pmm", "MNAR", "MNAR_pmm"))
```

@@

Data are simulated and included in the `micemd` package. We will use the following variables:

- `region` Cluster variable,
- `gender` Gender (0=male, 1=female),
- `age` Age of the patient,
- `height` Height in meters,
- `weight` Weight in kilograms,
- `time`` Response time in minutes (inclusion-restriction variable).

Load the data:
```{r obesity}
data("data_heckman", package = "micemd")
```

Load the data:
```{r}
#data("obesity", package = "miceheckman")
```

Select relevant variables:
```{r bmi}
#dat <- select(obesity, -"bmi")
```

Create predictor matrix:
```{r}
#pred <- quickpred(dat)   # predictor matrix
#pred["weight","cluster"]<- -2 # clustering variable
#pred["weight","rt"] <- -3 #inclusion-restriction variable
#plot_pred(pred)
```

Set the Heckman model as imputation method:
```{r}
#meth <- make.method(dat) # methods vector
#meth["weight"]<-"2l.heckman"
```

Visualize missing data pattern:
```{r obesity-md}
plot_pattern(dat)
```

Create predictor matrix:
```{r obesity-pred, eval = FALSE}
pred <- quickpred(dat)   # predictor matrix
pred[,"Cluster"] <- -2 # clustering variable
pred[, "Weight"] <- -3 #inclusion-restriction variable
plot_pred(pred)
```

Set the Heckman model as imputation method:
```{r, eval = FALSE}
meth <- make.method(dat) # methods vector
meth["weight"]<-"2l.heckman"
```

Impute the missingness:
```{r eval=FALSE, cache=TRUE}
imp <- mice(
  dat, # dataset with missing values
  m = 5, # number of imputations
  maxit = 10,
  meth = meth, #imputation method vector
  pred = pred, #imputation predictors matrix
  meta_method = "reml",
  printFlag = FALSE
)
```

# Discussion

@@

>>>>

```{r, include=FALSE, eval=TRUE}
library(here)
source(here("heckman","ipdfunction.R"),local =knitr::knit_global())
load(here("heckman","Obesity.rda"))
```

@@

```{r, dpi = 300,out.width="70%", fig.cap="Missing pattern", eval=TRUE}

#ggmice::plot_pattern(data=Obesity,cluster="Cluster") does not work!! bug
#ggmice::plot_pattern(data=Obesity,cluster=Obesity$Cluster) does not work!! bug
library(ggpubr)
myplots <- lapply(1:5, function(i) {
  ggmice::plot_pattern(setDT(Obesity)[Cluster==i])+
  ggplot2::ggtitle(paste0("Cluster", i))
 })
ggarrange(myplots[[1]], myplots[[3]], nrow=1,common.legend = TRUE, legend="bottom")
```

@@

```{r obesity-icc, eval=TRUE}
Nulmodel <- lme4::lmer(BMI ~ 1 + (1|Cluster), data = Obesity)
performance::icc(Nulmodel)
```

@@

```{r obesitymeth, eval=TRUE}
library(micemd)
meth_mar <- micemd::find.defaultMethod(Obesity, ind.clust=1, I.small = 7,
                               ni.small = 100, prop.small = 0.4)
meth_mar["BMI"]<- "~ I(Weight / (Height)^2)"
meth_mar["Age"]<-"2l.2stage.pmm" 
```

For these imputation models, it is necessary to specify the prediction matrix, with the cluster
variable labelled as -2 and the predictor variable measured within clusters labelled as 2, encompassing all variables. We need to suprime the variable Time as this variable is not specified in the main model. We also  modify the relationship between BMI, weight and height in the prediction matrix to avoid circular predictions. Then we proceed to run the imputation model.
```{r obesity-predmar,out.width="70%", eval=TRUE}
pred_mar <- mice(Obesity, maxit = 0)$pred 
pred_mar[,"Cluster"] <- -2 # clustering variable
pred_mar[,"Time"] <- 0
pred_mar[pred_mar==1] <- 2
pred_mar[c("Height", "Weight"), "BMI"] <- 0
ggmice::plot_pred(pred_mar)
imp_mar <- mice::mice(data = Obesity, meth = meth_mar, pred = pred_mar,
                m=10, seed = 123, printFlag = FALSE)
```

```{r obesity-predmar1,eval=TRUE}
summary(complete(imp_mar,"long")$Weight)
```
We are also contemplating the utilisation of the ???pmm??? option, as the values imputed using a fully parametric method may be implausibly low for some patients.

```{r obesity-predmar_pmm, eval=TRUE}
meth_mar["Weight"]<-"2l.2stage.pmm" 
imp_mar_pmm <- mice(data = Obesity, meth = meth_mar, pred = pred_mar,
                    m=10, seed = 123, printFlag = FALSE)
```

```{r obesity-predmar_pmm1,eval=TRUE}
summary(complete(imp_mar_pmm,"long")$Weight)
ggmice::plot_trace(imp_mar_pmm, "Weight")
```

@@

```{r obesity-predmnar, eval=TRUE}
pred_mnar <- pred_mar
pred_mnar["Weight","Time"]<- -3
ggmice::plot_pred(pred_mnar)
```

@@

```{r obesity-predmnar1, eval=TRUE}
meth_mnar <- meth_mar
meth_mnar["Weight"]<- "3l.2stage.heckman"
```
Then we proceed to run the imputation model as before, after executing these imputation procedures, it is essential to assess convergence and the coherence of the imputed values.
```{r obesity-predmnar2,eval=TRUE}
imp_mnar<- mice(data = Obesity, meth = meth_mnar, pred = pred_mnar,
                m=10, seed = 123, printFlag = FALSE)
summary(complete(imp_mnar,"long")$Weight)
```

@@

```{r obesity-predmnar3,eval=TRUE}
summary(complete(imp_mnar,"long")$Weight)
```
Consequently, as before we  use the 'pmm', option but this time for the Heckman imputation, this approach ensures that the imputed values remain within the range of observable values. We then run the imputation model but this time using the option of pmm, to assure that weight values are in the range of the observable data, this can be implemented by setting the pmm parameter to true.

```{r obesity-predmnarp, eval=TRUE}
imp_mnar_pmm <- mice(data = Obesity, meth = meth_mnar, pred = pred_mnar,
                     m=10, seed = 123, pmm = T,  printFlag = FALSE)
```

We check the convergency of the results
```{r obesity-predmnarp1, eval=TRUE}
summary(complete(imp_mnar_pmm,"long")$Weight)
ggmice::plot_trace(imp_mar_pmm, "Weight")
```

@@

```{r models, eval=TRUE}
library(ggplot2)
cc_rs<- with(setDT(Obesity)[complete.cases(Obesity),],
             lme( BMI ~ Age + FamOb + Gender, random=~1+Age|Cluster))
mar_rs <- with(imp_mar,lme( BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
mar_pmm_rs <- with(imp_mar_pmm,lme( BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
mnar_rs<- with(imp_mnar,lme(BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
mnar_pmm_rs<- with(imp_mnar_pmm, lme(BMI ~ Age + FamOb + Gender,random=~1+Age|Cluster))
list_models<-list(cc_rs,mar_rs,mar_pmm_rs,mnar_rs,mnar_pmm_rs)
plot_models(list_models,
            mod_name = c("Complete case", "MAR","MAR_pmm", "MNAR", "MNAR_pmm"))
```

@@

@@

# Conclusion
